Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-03-30 RMS/responses.tex
almost finished, ready to be final version of revision

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Corrections to Manuscript No. 15-TIE-3480}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

%\section{Answers to the Reviewers}
\section{Reponses to Reviewer 1}
\subsection{General Comments}
\begin{itemize}

@@

\ANSWER Thanks for the pros. We are happy that the work in the paper
interests the readers. We hope that it provides a new and formal way
to verify a real-time system, which is able to investigate important
problems (such as periodic task scheduling) from a more realistic
aspect.

@@

\ANSWER Thanks for the suggestion. As mentioned in the paper, the
system under verification is a real-world RMS implementation, which is
used in an avionic control system. The scenarios presented in Section
V.B are also real scenarios from the online system. Our industrial
partner has very strict requirements in the timing. They use at most
$5$ tasks for scheduling to avoid high overhead.

On the other hand, ``larger'' scenarios are not the same as
``complex'' scenarios in our model, as discussed in the revised
Section V.B. We have examined our approach by verifying randomly
generated scenarios where the number $n$ of tasks are over $10$. In
that case, the least common multiple of the periods of all tasks can
be as large as $10^5$ times the interrupt cycle $T$. All feasible
combinations of execution traces are checked in our approach. Our
industrial partner agrees that it satisfies their practical
requirements. Furthermore, the number $n$ of tasks can be $20$ and
even more if we allow different tasks possess a same period.

@@

\ANSWER We are happy that the readers are interested in the detailed
and technical proof, which is now included in the Appendix. It is true
that readers may be interested in the detailed proof of Theorem~2 to
feel convinced of the completeness of our approach. However, the proof
needs a bit more theoretical background about rewriting logic and more
technical details of our model, which we have simplified in the paper
in order to ease the understanding of our approach. We would prefer to
refer the detailed proof to a non-anonymous technical report, if
possible, in the final version.

@@

\ANSWER Thank you for the invaluable suggestions. Section IV has been
revised. Some unnecessary detailed definitions presented as code are
removed, and replaced by more explanations. We hope that the current
presentation would make Section IV more easy-to-read, even for an
audience who did not know Real-Time Maude before. On the other hand,
Section VI (Related Work) is also improved. We compare our work with
the existing theoretical work and the existing verification work in a
more complete way.

@@

\ANSWER Thanks for the pointing this out. The related work section is
now improved. More comparisons with the theoretical approaches and
with [24,25] are added. The differences with [24,25] in the models and
in the way of modeling are mainly discussed. On the other hand, it is
a bit difficult to compare the efficiency and scalability between
[24,25] and our approach. One reason is that we have a different
setting with [24] and [25], because the objectives are different: we
aim at verifying an RMS implemention from a real-world real-time
system, while [24] and [25] targeted the RMS algorithm itself. Another
reason is that [24] and [25] presented no discussion on the efficiency
and scalability. Furthermore, they used \verb|TMSVL| and an extension
of SPIN, respectively, as verification tools, which are not
open-source, making us unable to re-implement their work.

@@

\ANSWER Sorry for that. The statement has been corrected. The work
presented in the paper aims at modeling and verifying a realistic RMS
implementation in an industrial avionic system. It is a piece of real
work in the industry. Even though we believe that our approach could
be applied to other similar systems as well, it means more to that
specific one. In this sense, ``sufficient'' means that the details
described in our model meet the requirements and expectations from the
users and the engineers.

@@

To us, there is a trade off between these two directions. We did
choose the latter according to the following considerations. By our
experience, \emph{the development engineers (who write the code) and
  verification engineers (who verify the code or the system) are
  usually not the same persons.} When a verification engineer models a
system, he is actually abstacting the system or the codes. It is
difficult to ensure that the model behaves the same as the system or
the codes.  And it is more difficult to make other engineers and users
believe that the model behaves the same as the system or the
codes. Furthermore, the developers and users hope that the model
behaves not only the same as what they had in mind, but also the same
as what is written in the code. A way out is to make the model not
only \emph{behave like} the code, but also \emph{``look'' like} the
code. This requires the model to be as accurate as possible, based on
a fact that the expressivity of the modeling language is powerful
enough. It makes the model specific, but improves the engineers'
confidence in the model and the verification results. That is also why
we emphasize the correspondences between the functions in the model
and the lines in the pseudocode, when we introduce our model in
Section~IV.

@@

\ANSWER Thanks for this comment. The section about soundness and
completeness is now made more self-contained. The detailed proof of
Theorem~2 is included in the Appendix. On the other hand, since we
chose the accurate way to achieve the model as discussed above, yes,
the proof is a specific one.

@@

\ANSWER Sorry for the misleading in the text. In fact, \verb|bottom|
and \verb|#| are constructors of sort \verb|Stack|. This part is
modified now. Some detailed definitions in code are removed, and are
replaced by more explanations. We hope that the current presentation
would make Section IV more easy-to-read.

@@

\ANSWER Sorry for this incompleteness. \verb|NzNat| did mean non-zero
natural numbers. \verb|NzNat| is now replaced by \verb|Nat| to reduce
unnecessary complexity for the readers.

@@

\ANSWER Thank you for the insightful comment. We agree that RMS has
been intensively studied, because of its importance. However, most of
the results are based on the assumption that \emph{task switching does
  not take time}. That assumption is not realistic. Models with and
without the assumption have totally different behaviors. For instance,
as shown in Figure~2(b), in the RMS implementation we verified, the
switching from time $9$ to $11$ blocks the interrupt handling, hence
delaying the initiation time of $\tau_1$. To our knowledge, this kind
of phenomenon would not happen in the theoretical analysis
models. Although there is very few results that take into account the
switching overhead (as discussed in Section VI), none of them is
generic and our system under verification is not in their scope.

@@

\hide{
is a very simple scheduling and is the optimum at the same time among
the fixed-prority ones, making it widely applied and sufficiently
investigated. Besides the ideal setting, the schedulability of RMS was
also investigated under different kinds of implementations with
consideration about overhead, as we discuss in the related work
section (Section VI). Despite the reason that the existing results do
not fit our target system using RMS, the most important motivation to
develop the work presented is that, we aim at verifying the system not
only for the schedulability, but also for the correctness. That is, we
want to verify that the target system does schedule the tasks under
the RMS algorithm. The work presented should be seen as a verification
problem, instead of a schedulability test.
}

@@

\ANSWER Thank you for the question. When an automatic formal
verification tool (such as model checker) is used to verify a given
system, it explores the whole system state space to check whether the
given property holds. Then the state space explosion is the most
common problem for an automatic verification tool. That means the
state space of the system/model is too huge for the tool to explore
with acceptable temporal and spatial cost. The scale of the state
space depends on the complexity of the model of the system.

@@

On the other hand, the assumptions are not so rigid. For example, the
mask mechanism considered makes the job initiation times of task
$\tau_i$ possibly not equal $kT_i$, implying that the deadlines of the
jobs may be not equal to $(k+1)T_i$ (see assumption A1'). This is also
discussed in the beginning of Section IV, pointing out the blocking at
time $10$ in Figure~2(b). Another example is that non-determinism
exists in our model, as discussed in Section~VI.

>>>>

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
%\title{Corrections to Manuscript No. 15-TIE-3480}
\title{Responses to Reviews}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

\section{General Response}
We thank all the reviewers for their careful considerations on our
paper. The comments are very insightful and invaluable. Thanks to the
comments, we have improved the paper a lot to make it clearer and
better presented.

The main changes in the paper are as following:
\begin{itemize}
\item
Section~IV about the formal model of the target implementation is
revised. Some technical details have been simplified and more literal
explanations are used to deliver our ideas.

\item
Section~V.B is expanded. Some discussion on the efficiency of our
approach is added.

\item 
The detailed technical proof of Theorem~2 in Section~V.C has been
included in the Appendix.

\item
Section~VI about related work has been improved. 
\end{itemize}

Detailed responses to all the comments are shown in the rest of this
document.

%\section{Answers to the Reviewers}
\section{Responses to Reviewer 1}
\subsection{General Comments}
\begin{itemize}

@@

\ANSWER Thanks for the pros. We are happy that the work in the paper
interests the readers. We hope that it provides a new and formal way
to verify a real-time system, which is able to investigate important
problems (such as periodic task scheduling) in a more realistic level.

@@

\ANSWER Thanks for the suggestions. As mentioned in the paper, the
system under verification is a real-world RMS implementation, which is
used in an avionic control system. The scenarios presented in Section
V.B are also real scenarios from the online system. Our industrial
partner has very strict timing requirements. They use at most 5 tasks
for scheduling to avoid high overhead. They agree that the
verification results presented in the paper satisfy their practical
requirements.

On the other hand, as presented in the revised Section V.B, we have
examined our approach by verifying randomly generated scenarios where
the number $n$ of tasks are over $10$. In that case, the least common
multiple of the periods of all tasks can be as large as $10^5$ times
the interrupt cycle $T$. All feasible combinations of execution traces
are checked in our approach. Furthermore, the number $n$ of tasks can
be $20$ and even more if we allow different tasks possess a same
period.

@@

\ANSWER We are happy that the readers are interested in the detailed
technical proof, and we have now included it in the Appendix. It is
true that readers may be interested in the detailed proof of Theorem~2
to feel convinced of the completeness of our approach. However, the
proof needs a bit more theoretical background about rewriting logic
and more technical details of our model, which we have simplified in
the paper in order to ease the understanding of our approach. We would
prefer to refer the detailed proof to a non-anonymous technical
report, if possible, in the final version.

@@

\ANSWER Thank you for the invaluable suggestions. Section IV has been
revised. Some unnecessary detailed definitions presented as code are
removed, and replaced by more literal explanations. We hope that the
current presentation would make Section IV more easy-to-read, even for
an audience who did not know Real-Time Maude before. On the other
hand, Section VI (Related Work) is also improved. We compare our work
with the existing theoretical work and the existing verification work
in a more complete way.

@@

\ANSWER Thanks for the pointing this out. The related work section is
now improved. More comparisons with the theoretical approaches and
with [24,25] are added. The differences with [24,25] in the models and
in the way of modeling are mainly discussed. On the other hand, it is
a bit difficult to compare the efficiency and scalability between
[24,25] and our approach. One reason is that we have a different
setting with [24] and [25], since the objectives are different: we aim
at verifying an RMS implementation from a real-world real-time system,
while [24] and [25] targeted the RMS algorithm itself. Another reason
is that [24] and [25] presented no discussion on the efficiency and
scalability. Furthermore, they used \verb|TMSVL| and an extension of
SPIN, respectively, as verification tools, which are not open-source,
making us unable to re-implement their work.

@@

\ANSWER Sorry for the confusion. The statement has been corrected. The
work presented in the paper aims at modeling and verifying a realistic
RMS implementation in an industrial avionic system. It is a piece of
real work in the industry. In this sense, ``sufficient'' means that
the details described in our model meet the requirements and
expectations from the users and the engineers. On the other hand, we
believe that our approach could be applied to other similar systems as
well.

@@

We chose the latter according to the following considerations. By
experience, \emph{the development engineers (who write the code) and
  verification engineers (who verify the code or the system) are
  usually not the same persons.} When a verification engineer models a
system, he is actually abstracting the system or the code. It is
difficult to ensure that the model behaves the same as the system or
the code.  And it is more difficult to make other engineers and users
believe that the model behaves the same as the system or the
code. Furthermore, the developers and users hope that the model
behaves not only the same as what they had in mind, but also the same
as what is written in the code. A way out is to make the model not
only \emph{behave like} the code, but also \emph{``look'' like} the
code. This requires the model to be as accurate as possible, based on
a fact that the expressiveness of the modeling language is powerful
enough. It makes the model specific, but improves the engineers'
confidence in the model and the verification results. That is also why
we emphasize the corresponding relationship between the functions in
the model and the lines in the pseudocode, when we introduce our model
in Section~IV.

@@

\ANSWER Thanks for this comment. The detailed proof of Theorem~2 is
now included in the Appendix. On the other hand, since we chose the
accurate way to achieve the model as discussed above, yes, the proof
is a specific one.

@@

\ANSWER Sorry for the misleading presentation in the text. In fact,
\verb|bottom| and \verb|#| are constructors of sort \verb|Stack|. This
part is modified now. Some detailed definitions in code are removed,
and are replaced by more explanations. We hope that the current
presentation would make Section IV more easy-to-read.

@@

\ANSWER Sorry for this incompleteness. \verb|NzNat| meant non-zero
natural numbers. \verb|NzNat| is now replaced by \verb|Nat| to reduce
unnecessary complexity for the readers.

@@

\ANSWER Thank you for the insightful comment. We agree that RMS has
been intensively studied because of its importance. However, most of
the results are based on the assumption that \emph{task switching does
  not take time}. That assumption is not realistic. Models with and
without the assumption have totally different behaviors. For instance,
as shown in Figure~2(b), in the RMS implementation we verified, the
switching from time $9$ to $11$ blocks the interrupt handling, hence
delaying the initiation time of $\tau_1$. To our knowledge, this kind
of phenomenon would not happen in the theoretical analysis
models. Although there is very few results that take into account the
switching overhead (as discussed in Section VI), none of them is
generic and our system under verification is not in their scope.

@@

\hide{
is a very simple scheduling and is the optimum at the same time among
the fixed-priority ones, making it widely applied and sufficiently
investigated. Besides the ideal setting, the schedulability of RMS was
also investigated under different kinds of implementations with
consideration about overhead, as we discuss in the related work
section (Section VI). Despite the reason that the existing results do
not fit our target system using RMS, the most important motivation to
develop the work presented is that, we aim at verifying the system not
only for the schedulability, but also for the correctness. That is, we
want to verify that the target system does schedule the tasks under
the RMS algorithm. The work presented should be seen as a verification
problem, instead of a schedulability test.
}

@@

\ANSWER Thank you for the question. When model checking technique is
used to verify a given system, the tool (model checker) explores the
whole system state space to check whether the given property
holds. All feasible combinations of execution traces are checked. As a
result, the state space explosion is the most common problem for a
model checker. That means the state space of the system/model is too
huge for the tool to explore with acceptable temporal and spatial
cost. The scale of the state space depends on the complexity of the
model of the system.

@@

On the other hand, the assumptions are actually not that rigid. For
example, the mask mechanism considered makes the job initiation times
of task $\tau_i$ possibly not equal $kT_i$, implying that the
deadlines of the jobs may be not equal to $(k+1)T_i$ (see assumption
A1'). This is also discussed in the beginning of Section IV, pointing
out the blocking at time $10$ in Figure~2(b). Another example is that
non-determinism exists in our model, as discussed in Section~VI.

### 2016-03-30 RMS/submission.tex
almost finished, ready to be final version of revision

We have analyzed our model in ten different scenarios, including both
realistic ones provided by our industrial partner and experimental
ones designed by ourselves, four of them are described below:
\begin{itemize}
\item Scenario~(i) with two tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=3ms$, $T_2=25ms$ and $C_2=7ms$.
\item Scenario~(ii) with two tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=2ms$, $T_2=25ms$ and $C_2=2.3ms$.
\item Scenario~(iii) with three tasks $\tau_1$, $\tau_2$ and $\tau_3$:
  $T_1=5ms$, $C_1=2.7ms$, $T_2=10ms$, $C_2=2ms$, $T_3=25ms$ and
  $C_3=3ms$.
\item Scenario~(iv) with three tasks $\tau_1$, $\tau_2$ and $\tau_3$:
  $T_1=5ms$, $C_1=2.5ms$, $T_2=10ms$, $C_2=1.5ms$, $T_3=15ms$ and
$C_3=4.5ms$.
\end{itemize}
Note that thanks to the powerful expressivity of Real-Time Maude, we
only need to define an initial state of sort \verb|System| to specify
a given task set. No necessity to modify the model is needed.

@@

We also randomly generate test scenarios under the above setting to
exam the efficiency of our approach for verifying the schedulability.
Among the $50$ generated test scenarios with $5$ tasks, we find that
the execution time of the model checking command for each scenario
varies from about $300ms$ up to a timeout, which is set to $90$
minutes. This is because the efficiency of model checking depends on
the scale of the state space. And the scale is further positively
correlated with $mn$, where $m$ is the upper bound of $timer$ and $n$
is the number of periodic tasks. By the generated test scenarios, it
turns out that the model checking command for schedulability in our
model is able to handle scenarios, where $mn$ is up to about $10^6$,
in an acceptable period of time say $90$ minutes.

@@

Time-robustness is a set of properties expected from a well-behaved
real-time rewrite theory. We avoid introducing the accurate definition
of time-robustness, but instead give the following lemma to prove
time-robustness:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:timerobustness}
Let $\cal R$ be an object-oriented specification with a standard tick
rule, and let the infinity element \verb|INF| be the only element in
the time domain which is not a normal time value.
%and let the time domain be \emph{linear}.
Then $\cal R$ is time-robust if the following conditions are satisfied
for all appropriate ground terms $t$ and $r,r'$:
\begin{enumerate}
\item
[(i)] \verb|mte(delta(|$t$\verb|,|$r$\verb|))| $=$
\verb|mte(|$t$\verb|) monus |$r$, for all $r\le$
\verb|mte(|$t$\verb|)|;

>>>>

We have analyzed our model in ten different scenarios, including both
realistic ones provided by our industrial partner and experimental
ones designed by ourselves, four of them are described below:
\begin{itemize}
\item Scenario~(i) with two tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=3ms$, $T_2=25ms$ and $C_2=7ms$.
\item Scenario~(ii) with two tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=2ms$, $T_2=25ms$ and $C_2=2.3ms$.
\item Scenario~(iii) with three tasks $\tau_1$, $\tau_2$ and $\tau_3$:
  $T_1=5ms$, $C_1=2.7ms$, $T_2=10ms$, $C_2=2ms$, $T_3=25ms$ and
  $C_3=3ms$.
\item Scenario~(iv) with three tasks $\tau_1$, $\tau_2$ and $\tau_3$:
  $T_1=5ms$, $C_1=2.5ms$, $T_2=10ms$, $C_2=1.5ms$, $T_3=15ms$ and
$C_3=4.5ms$.
\end{itemize}
Note that thanks to the powerful expressiveness of Real-Time Maude, we
only need to define an initial state of sort \verb|System| to specify
a given task set. No necessity to modify the model is needed.

@@

The results above have demonstrated that our approach is capable of
handling realistic industrial systems. However to further exam the
efficiency of our approach, we also apply our method to larger test
scenarios. Test scenarios are generated randomly. We verify the
schedulability of them under the above setting on an Intel Core 2 Quad
Q9550, 2.83GHz, 4-cores machine with 8GB RAM running 64-bit Ubuntu
15.04. Among the 50 generated test scenarios with 5 tasks, we find
that the execution time of the model checking command for each
scenario varies from about $300ms$ up to a timeout, which is set to
$90$ minutes. This is because the efficiency of model checking depends
on the scale of the state space. And the scale is further positively
correlated with $mn$, where $m$ is the upper bound of $timer$ and $n$
is the number of periodic tasks. By the generated test scenarios, it
turns out that the model checking command for schedulability in our
model is able to handle scenarios, where $mn$ is up to about $10^6$,
in an acceptable period of time say $90$ minutes.

@@

Time-robustness is a set of properties expected from a well-behaved
real-time rewrite theory. We avoid introducing the accurate definition
of time-robustness, but instead give the following lemma to prove
time-robustness:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:timerobustness}
Let $\cal R$ be an object-oriented specification with a standard tick
rule, and let the infinity element \verb|INF| be the only element in
the time domain which is not a normal time value.
%and let the time domain be \emph{linear}.
Then $\cal R$ is time-robust if the following conditions are satisfied
for all appropriate ground terms $t$ and $r,r'$:
\begin{enumerate}
\item
[(i)] \verb|mte(delta(|$t$\verb|,|$r$\verb|))| $=$
\verb|mte(|$t$\verb|) monus |$r$, for all $r\le$
\verb|mte(|$t$\verb|)|, where \verb|monus| is the minus operation
defined on sort \verb|Time|;
